Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-21 sections/lang.tex
more semantics

\[
    \frac{}{\langle\finchliteral(val), (e, t)\rangle \rightarrow val} \quad
    \frac{}{\langle\finchvalue(ex, type), (e, t)\rangle \Rightarrow \langle ex, t \rangle} \quad
\]
\[
    \frac{}{\langle\finchvar(name), (e, t)\rangle \rightarrow e(\finchvar(name))} \quad
    \frac{}{\langle\finchindex(name), (e, t)\rangle \rightarrow e(\finchindex(name))}
\]
\[
    \frac{\langle args_i, (e, t) \rangle \Rightarrow vals_i \quad \langle f, (e, t) \rangle \Rightarrow g}
    {\langle\finchcall(f, args...), (e, t)\rangle \rightarrow \langle g(vals...), t \rangle}
\]
\[
\frac{\langle head, s_1 \rangle \rightarrow s_2}
{\langle\finchblock(head, tail...), s_1 \rangle \rightarrow \langle \finchblock(tail...), s_2 \rangle} \quad
\frac{}{\langle\finchblock(), s\rangle \rightarrow s}
\]
\[
\frac{\langle cond, s \rangle \Rightarrow true}
{\langle\finchsieve(cond, body), s\rangle \rightarrow \langle body, s \rangle}\quad
\frac{\langle cond, s \rangle \Rightarrow false}
{\langle\finchsieve(cond, body), s\rangle \rightarrow s}
\]
\[
\frac{\langle val, (e, t) \rangle \rightarrow val'}
{\langle\finchdefine(var, val, body), (e, t)\rangle \rightarrow \langle body, (e[var \mapsto val'], t) \rangle}
\]
\[
\frac{s = (e, t) \quad e(\finchmode(tns)) = read \quad e(tns) = tns' \quad  \langle init, s \rangle \rightarrow init' \quad  \langle dims_i, s \rangle \rightarrow dims_i'}
{\langle\finchdeclare(tns, init, dims...), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto update], \langle declare(tns', init', dims'...), t \rangle)}
\]
\[
\frac{s = (e, t) \quad e(\finchmode(tns)) = read \quad e(tns) = tns'}
{\langle\finchthaw(tns, init, dims...), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto update], \langle thaw(tns'), t \rangle)}
\]
\[
\frac{s = (e, t) \quad e(\finchmode(tns)) = update \quad e(tns) = tns'}
{\langle\finchfreeze(tns, init, dims...), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto read], \langle freeze(tns'), t \rangle)}
\]

>>>>

\[
    \frac{}{\langle\finchliteral(val), (e, t)\rangle \rightarrow val} \quad
    \frac{}{\langle\finchvalue(ex, type), (e, t)\rangle \Rightarrow \langle ex, t \rangle} \quad
\]
\[
    \frac{}{\langle\finchvar(name), (e, t)\rangle \rightarrow e(\finchvar(name))} \quad
    \frac{}{\langle\finchindex(name), (e, t)\rangle \rightarrow e(\finchindex(name))}
\]
\[
    \frac{\langle args_i, (e, t) \rangle \Rightarrow vals_i \quad \langle f, (e, t) \rangle \Rightarrow g}
    {\langle\finchcall(f, args...), (e, t)\rangle \rightarrow \langle g(vals...), t \rangle}
\]
\[
\frac{\langle head, s_1 \rangle \rightarrow s_2}
{\langle\finchblock(head, tail...), s_1 \rangle \rightarrow \langle \finchblock(tail...), s_2 \rangle} \quad
\frac{}{\langle\finchblock(), s\rangle \rightarrow s}
\]
\[
\frac{\langle cond, s \rangle \Rightarrow true}
{\langle\finchsieve(cond, body), s\rangle \rightarrow \langle body, s \rangle}\quad
\frac{\langle cond, s \rangle \Rightarrow false}
{\langle\finchsieve(cond, body), s\rangle \rightarrow s}
\]
\[
\frac{\langle val, (e, t) \rangle \rightarrow val'}
{\langle\finchdefine(var, val, body), (e, t)\rangle \rightarrow \langle body, (e[var \mapsto val'], t) \rangle}
\]
\[
\frac{s = (e, t) \quad e(\finchmode(tns)) = read \quad e(tns) = tns' \quad  \langle init, s \rangle \rightarrow init' \quad  \langle dims_i, s \rangle \rightarrow dims_i'}
{\langle\finchdeclare(tns, init, dims...), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto update], \langle declare(tns', init', dims'...), t \rangle)}
\]
\[
\frac{s = (e, t) \quad e(\finchmode(tns)) = read \quad e(tns) = tns'}
{\langle\finchthaw(tns), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto update], \langle thaw(tns'), t \rangle)}
\]
\[
\frac{s = (e, t) \quad e(\finchmode(tns)) = update \quad e(tns) = tns'}
{\langle\finchfreeze(tns), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto read], \langle freeze(tns'), t \rangle)}
\]

$E$ is an evaluation context that applies anywhere in the syntax tree.

\[
\frac{e(tns) = tns' \langle unfurl(tns'), t\rangle \rightarrow \langle tns'', t'\rangle}
{\langle \finchloop(i, ext, E[\finchaccess(tns, j..., i)]), (e, t)\rangle \rightarrow \langle \finchloop(i, ext, E[\finchaccess(tns'', j..., i)]), (e, t')\rangle}
\]

\[
\frac{}
{\splitfrac{\langle \finchloop(i, ext, E[\finchaccess(sequence(phase(range, body), tail...), j..., i)]), s\rangle
    }{\rightarrow \langle \finchblock(\finchloop(i, ext \cup range, E[\finchaccess(body, j..., i)]), \finchloop(i, ext, E[\finchaccess(sequence(tail...), j..., i)])), s\rangle}}
\]

\[
\frac{}
{\langle \finchloop(i, ext, E[\finchaccess(sequence(), j..., i)]), s\rangle \rightarrow \langle \finchblock(), s\rangle}
\]

\[
\frac{}
{\splitfrac{\langle \finchloop(i, extent(a, b), E[\finchaccess(stepper(next, body), j..., i)]), s\rangle
    }{\rightarrow \langle \finchblock(\finchloop(i, ext \cup range, E[\finchaccess(body, j..., i)]), \finchloop(i, ext, E[\finchaccess(sequence(tail...), j..., i)])), s\rangle}}
\]
